### 2015-02-24 Paper/finalBLAU.tex
Einpflegen der final remarks

\documentclass[12pt,a4paper]{scrartcl}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[color=yellow!15]{todonotes}
\usepackage{graphicx}
\usepackage[numbers, sort&compress, square]{natbib}

\usepackage{subcaption}

\usepackage{titlesec}
\usepackage{color}
\usepackage{booktabs}

% used for line numbering
\usepackage{lineno}

\usepackage[colorlinks, linkcolor=black, citecolor=blue, urlcolor=blue, bookmarks]{hyperref}

\usepackage{geometry}
\geometry{a4paper, top=35mm, left=30mm, right=30mm, bottom=35mm,
headsep=10mm, footskip=12mm}

\title{\textbf{ \Large {Decentralized autonomous sensor fault detection in wireless sensor networks using neural networks}}}
\author
{\large Katrin Jahr, Robert Schlich\\
\\
\normalsize{Degree program “Civil Engineering” (M.Sc.)}\\
\normalsize{Bauhaus-Universität Weimar, Germany}\\
\\
\normalsize{katrin.jahr@uni-weimar.de \qquad}
\normalsize{robert.schlich@uni-weimar.de}
}
\date{}

\clubpenalty = 10000 % schliesst Schusterjungen aus
\widowpenalty = 10000 \displaywidowpenalty = 10000% schliesst Hurenkinder aus

\newenvironment{sciabstract}{%
\begin{quote} \itshape}
{\end{quote}}

\begin{document}

% Double-space the manuscript.

\baselineskip20pt

\frenchspacing

% Make the title.

\maketitle 
\vspace{-2em}

% 
\sloppy
\setlength{\emergencystretch}{6pt}
\hyphenation{sample-rate auto-nomous moni-toring methods during}

\begin{sciabstract}

The dependability and the accuracy of structural health moni\-toring systems can be affected by sensor faults. 
In this paper, the design and implementation of a wireless structural health monitoring system, capable of decentralized autonomous fault detection, are presented. 
For self-detecting sensor faults, each sensor node predicts expected sensor data and compares it to the measured sensor data. 
The predictions are computed using neural networks based on measured sensor data of adjacent sensor nodes.
In laboratory experiments, devised to validate the proposed approach, several simulated sensor faults are detected.
These results indicate that the use of neural networks for fault detection increases the dependability and the accuracy of structural health monitoring systems.

\end{sciabstract}

%----------------------------------------------------------------------------------------
%
%\linenumbers % Schaltet Zeilennummerierung ein
%\modulolinenumbers[5] % nur jede 5. Zeile
%\section*{Dictionary}
%
%\begin{tabular}{|l|l|}
%\hline 
%Sensorknoten (SunSPOT) & sensor node \\ 
%\hline 
%einzelner Messsensor (Thermometer) & sensor \\ 
%\hline 
%Knoten im neuronalen Netz & neuron \\ 
%\hline 
%eine abgeschlossene Messungreihe & test run \\ 
%\hline 
%gemessene Werte & sensor data \\ 
%\hline 
%vorhergesagte Werte & predicted data \\ 
%\hline 
%durch Vorhersage erwartete Werte & expected data \\ 
%\hline 
%einzelner Messwert & measurement \\ 
%\hline 
%Test & laboratory experiments\\ 
%\hline
%tatsächliche, nicht virtuelle Messung & actual measurement \\
%\hline
%Messaufbau & test setup \\
%\hline
%
%\end{tabular}

%----------------------------------------------------------------------------------------
\section*{Introduction}

Civil engineering structures are exposed to various external impacts during their lifetime. 
Structural health monitoring (SHM) systems can be deployed to evaluate the conditions and to ensure the structural stability of civil engineering structures.
\citet{BisbySHM} defines SHM as "a non-destructive \textit{in-situ} structural evaluation method that uses any of several sensors which are attached to, or embedded in, a structure".
\colorbox{cyan}{The} sensor data is collected by sensor nodes, and then analyzed and stored on a computer system. 
Through analyses of the sensor data, abnormal changes in material and geometric behavior can be \colorbox{cyan}{discovered} at an early stage.

Traditionally, the sensor nodes are connected to computer systems with cables.
The use of wired SHM systems has several disadvantages, for example expensive wiring, high installation and labor costs as well as inaccessibility of optimal sensor location with wires.

Eradicating wiring-specific problems, the sensor nodes in wireless SHM systems communicate via wireless transceivers \colorbox{cyan}{[...]}.
However, when using wireless SHM systems, special attention has to be paid to data transfer. 
To reduce transmission time, storage space and power consumption, and to decrease the probability of data loss, the transmitted data can be reduced using embedded computing.

As a data reduction strategy, the natural frequency of a structure can be calculated on the sensor nodes instead of transmitting all raw measurements.
Changes in the natural frequency result from changes in the vibration behavior, which indicate structural damage \citep{salawu1997detection}.
The natural frequency can be determined by analyzing the frequency spectrum of a vibration of a structure.
The frequency spectrum shows the relationship between the frequencies and magnitudes---magnitudes represent the influence a certain frequency has in a vibration.
The natural frequency is defined as the frequency with the highest magnitude.
While the frequencies are unique for a structure, the scale of the magnitudes depend not only on the structure, but also on the degree of the vibration and the sensor location at the structure.
During a vibration, the magnitudes on different locations are correlated with each other \citep{rao2011fast}.

Over their lifetime, the wireless sensor nodes can become inaccurate, faulty, or may even break.
A fault can be defined as a defect of a sensor that leads to an error. 
\colorbox{cyan}{An} error is the manifestation of a fault---an incorrect system state---that may result in a failure.
To ensure the dependability and the accuracy of the SHM system, sensor faults must be detected and isolated in real time. 
A well known approach to fault detection is the installation of physically redundant sensors.
Faulty sensors can be identified through the deviation of their measurements from the measurements of correlated sensors.
Physical redundancy, although efficient for sensor fault detection, causes increased installation and maintenance costs due to the multiple installation of sensors. 
Representing a more efficient approach, analytical redundancy typically uses mathematical functions mapping the characteristics of the structure and the correlations of the installed sensors \citep{Smarsly2014}. 
Specifically, virtual sensor measurements are computed for each sensor and then compared to the actual measurements. 
If the properties of a structure are known, finite element models can be used in combination with data from adjacent sensor nodes to calculate virtual measurements of a sensor, i.e. a priori knowledge about the structure is required.

Without a priori knowledge, analytical redundancy can be implemented into wireless sensor nodes based on artificial neural networks, that are used to predict measurements.
Artificial neural networks essentially consist of interconnected data processing units called neurons. 
The neurons are grouped in different layers; usually one input layer, a number of hidden layers, and one output layer.
Artificial neural networks are able to learn, which is achieved by adjusting the weights of the inter-neuron connections until a set of given input variables results in the desired output variables; a neural network can be trained to approximate any mathematical functions with any level of accuracy \citep{Li2011}.

In this study, a wireless SHM system capable of decentralized autonomous fault detection is presented.
The sensor nodes calculate the natural frequency of the structure to monitor the structural integrity.
The corresponding magnitudes are used for sensor fault detection:
A neural network is implemented into each sensor node. 
Using the magnitudes of past vibrations, each neural network is trained to map the correlation of the magnitudes calculated by different sensor nodes.

During runtime of the SHM system, the neural networks compute expected magnitudes for the respective sensor node.
Deviations between the expected and the measured magnitude indicates a sensor faults.

This paper is organized as follows:
First, a decentralized wireless structural health monitoring system is designed and implemented. 
Next, a neural network is implemented into each sensor node and trained to predict the sensor measurements of the specific node for detecting sensor faults in a decentralized manner. 
Finally, the system is tested in laboratory experiments \colorbox{cyan}{with simulated sensor faults} and the experimental results are discussed.

%----------------------------------------------------------------------------------------

\section*{Design and implementation of the wireless structural health monitoring system}
In this section, the wireless structural health monitoring system is introduced and the software implementation is described.
The wireless SHM system is composed of wireless sensor nodes and a host computer, both linked through a basestation.
The sensor nodes and the basestation are of type "Oracle Sun SPOT". 
The Sun SPOTs are equipped with several components, e.g. the sensor nodes include an internal accelerometer and eight independent RGB-LEDs.
The 3-axis digital output accelerometer ranges between $\pm$\,2\,g and $\pm$\,8\,g has a maximum sampling rate of 125\,Hz \citep{eDemo2010}.

The SHM system performs the following monitoring tasks, \colorbox{cyan}{as definded by \citet{BisbySHM}}:
1. data acquisition,
2. data processing,
3. data transmission, 
4. data storage,
5. diagnostics and 
6. information retrieval.
The data flow is illustrated in \autoref{fig:flow}.
Tasks 1 to 3, \colorbox{cyan}{representing a sampling event}, are executed by the sensor nodes: During system operation, the sensor nodes acquire acceleration measurements and determine the natural frequency as well as the corresponding magnitude of the structure using a fast Fourier transform (FFT) and a peak picking algorithm.
In every sensor node, the magnitudes of the natural frequencies are fed into a neural network to detect sensor faults.
The processed data is then transmitted wirelessly to the basestation and, finally, to the host computer.
On the host computer, task 4 is conducted by storing the data in a MySQL database.
Task 5 and task 6 are conducted on the host computer in further steps.

\begin{figure}[hbt]
    \centering
    \includegraphics[scale=0.9]{figures/dataflow_tasks_nn.eps}
    \caption{Data flow and associated SHM tasks}
    \label{fig:flow}
\end{figure}

The neural networks represent the structure of the SHM system.
Each sensor node of the SHM system embeds a distinct neural network. As shown in \autoref{fig:neuralnetwork}, a sensor node is represented by either an input neuron or an output neuron.
The measured magnitudes of other sensor nodes are fed into the input neurons of the neural network; the output neuron returns the predicted magnitudes of the \colorbox{cyan}{respective sensor node}.
Hidden neurons are used for computation and \colorbox{cyan}{are not represented physically}.

\begin{figure}[hbt]
    \centering
    \includegraphics{figures/neuralnetwork.pdf}
    \caption{Scheme of the neural network}
    \label{fig:neuralnetwork}
\end{figure}

The SHM system is implemented in an object-oriented way using Java programming language. 
Object orientation uses objects, which are instantiated using classes. 
\colorbox{cyan}{Restlicher Absatz verändert: }
A class includes methods and attributes.
Methods allow the objects to perform actions; attributes store object-specific data.
The SHM system consists of two packages, \texttt{sensornode} and \texttt{basestation}, that organize several Java classes.

\autoref{fig:UML-sn} illustrates the classes of the \texttt{sensor\-node} package in detail.
The package \texttt{sensor\-node} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Sample}, \texttt{Network\-Training}, \texttt{Communi\-cation} and \texttt{Main\-Spot}, forming the SHM system, and several classes obtained from the external Java library "SNIPE" \citep{SNIPE} which is used to implement the neural networks.
The classes \colorbox{cyan}{[...]} are embedded directly into the sensor nodes.

For starting the operation of the sensor node, the entry point of the program is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class, the SNIPE neural network classes, the \texttt{Network\-Training} class and the \texttt{Communi\-cation} class are created to autonomously perform the monitoring tasks.
The \texttt{Acceleration\-Sampler} class is responsible for measuring the acceleration.
There are two phases: First, \colorbox{cyan}{before the excitation,} the acceleration is measured with a low sampling rate.
Once the acceleration exceeds a pre-defined threshold, the second phase is entered by increasing the sampling rate. 
The measured values are stored in the sensor nodes in an array.
The different phases are indicated by lighting different LEDs of the sensor nodes.
The \texttt{FFT} class performs a FFT on the measured accelerations to transform the measured values from time domain into frequency domain. 
With the transformed data, the frequencies and the magnitudes of the measured vibrations are calculated.
Finally, the natural frequency is determined by extracting the maximum magnitude.
The \texttt{Sample} class defines objects representing analyzed measurements.
\colorbox{cyan}{The \texttt{Sample} objects store} the following parameters: The address of the sensor node, the maximum magnitude, the natural frequency, and a boolean value that indicates if a sensor fault was detected during runtime.
Using the SNIPE library, a neural network is implemented for fault detection.
The \texttt{Network\-Training} class defines a routine to train the neural network with \texttt{Sample} objects of previous sampling events.
The \texttt{Communi\-cation} class ensures the wireless communication by opening a radio connection between the sensor node and the basestation to communicate the \texttt{Sample} objects from the sensor node to the basestation.

\begin{figure}[htbp]
    \centering
    \includegraphics[width = \textwidth]{figures/uml-sensornode.pdf}
    \caption{Class diagram of the \texttt{sensornode} package}
    \label{fig:UML-sn}
\end{figure}

\autoref{fig:UML-bs} illustrates the classes of the \texttt{base\-station} package.
The package \texttt{base\-station} runs on the host computer and operates the basestation.
It includes the classes \texttt{Database\-Handler}, \texttt{Main\-Base}, and \colorbox{cyan}{\texttt{Sample}, as described in the package \texttt{sensornode}}.

The entry point of the program is the \texttt{run()} method in the \texttt{Main\-Base} class. 
The \texttt{Main\-Base} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.
The \texttt{Database\-Handler} class establishes a connection to a MySQL database installed on the host computer, creates a database table, unless the specified table already exists, and inserts the data into the database table.

\begin{figure}[htbp]
    \centering
    \includegraphics[width = \textwidth]{figures/uml-basestation.pdf}
    \caption{Class diagram of the \texttt{basestation} package}
    \label{fig:UML-bs}
\end{figure}

\section*{Laboratory experiments}
In this section, laboratory experiments, devised to validate the proposed SHM system, are described.
First, a description of the test setup is given; second, the data acquisition, data processing and sensor fault detection are illustrated; and, finally, the results are discussed.

To validate the proposed approach in laboratory experiments, the wireless SHM system is installed on a test structure, as shown in \autoref{fig:teststructure}.
The test structure is a 4-story shear-frame structure consisting of four steel plates of 25\,cm\,$\times$\,50\,cm\,$\times$\,0.8\,mm.
The plates are mounted on threaded rods with a vertical clearance of 23\,cm.
At the bottom of the structure, the rods are embedded into a solid block of 40\,cm\,$\times$\,60\,cm\,$\times$\,30\,cm.
The SHM system is installed on the test structure by mounting one wireless sensor node to each of the three top stories.
Ensuring a free vibration in natural frequency with little interferences, the structure is excited by a deflection of the top of the structure.

\begin{figure}[htb]

	\centering

	\begin{subfigure}{0.48\textwidth}
		\includegraphics[width=\textwidth]{figures/testsetupcorrect.jpg}
    	\caption{Regular sensor arrangement}
		\label{fig:teststructure}
    \end{subfigure}
    \quad
  	\begin{subfigure}{0.48\textwidth}
		\includegraphics[width=\textwidth]{figures/testsetuperror.jpg}
    	\caption{Simulated sensor fault}
        \label{fig:sensorfault}
	\end{subfigure}
  	\caption{Laboratory setup}
\end{figure}

Several test runs have been conducted.
During each test run, all sensor nodes run simultaneously.
A test run includes 1. a training phase and 2. a data collection phase \colorbox{cyan}{on every sensor node}.
The training phase consists of the implementation and the training of a neural network.
During the training of the neural network, several sampling events are performed and used as training input.
A sampling event includes the excitation of the structure, sampling of 512 acceleration measurements, calculation of the natural frequency and corresponding magnitude, and data exchange with the other sensor nodes.

The data collection phase \colorbox{cyan}{consists of any desired number} of sampling events, sensor fault detection, and data storage.
After every sampling event, the measured magnitudes of the other sensor nodes are used as input of the neural network to predict the expected magnitude of the embedding sensor nodes for sensor fault detection.
The deviation of the expected magnitude and the measured magnitude is calculated by the sensor node.
If the deviation exceeds a threshold, a sensor fault is assumed.
Upon sensor fault detection, the following variables are stored into the database: the time of the sampling event, the natural frequency, the measured magnitude, the expected magnitude, the deviation,  and a boolean value that indicates if a sensor fault has been detected.

The neural networks have been tested with varying numbers of sampling events \colorbox{cyan}{as training input} (training events) (\autoref{fig:traininputs}).
With increasing numbers of (training events), the prediction accuracy of the expected magnitude increases.
With 24 sampling events, a minimum deviation between the expected magnitude and the measured magnitude of 0.47\% in average has been reached.

\begin{figure}[htb]
    \centering
    \includegraphics{figures/traininputs.pdf}
    \caption{Correlation of number of training events and mean deviation between expected magnitudes and measured magnitudes}
    \label{fig:traininputs}
\end{figure}

To validate the autonomous sensor fault detection, a sensor fault has been simulated by shifting and rotating one of the sensor nodes after completing the training phase.
As a result, it has been observed that the expected magnitudes significantly deviate from the measured magnitudes (\autoref{tab:sensorfault}).
The SHM system has recognized the deviation, corroborating that the implemented sensor fault detection increases the dependability of the SHM system.

\begin{table}[hb]
	\centering
	\caption{Deviation between measured magnitudes and expected magnitudes with and without simulated sensor faults}
	\begin{tabular}{c c c c}
		\toprule
		Sensor fault & Measured magnitude & Expected magnitude & Deviation [\%]\\ 
		\midrule
		  & 2.1502 & 2.1389 & 0.5292 \\
		FALSE & 2.0817 & 2.0890 & 0.3529 \\
		  & 1.8663 & 1.8631 & 0.1726 \\
		\midrule	
		 & 1.8161 & 2.0725 & 12.3715\\
		TRUE & 1.6665 & 1.9404 & 14.1156\\
		 & 1.8238 & 2.1687 & 15.9035\\
		\bottomrule
	\end{tabular}
	\label{tab:sensorfault}
\end{table}

In total, through the embedded FFT analysis of acceleration measurements, the data traffic is decreased from 512 values to 4 values per sampling event and sensor node.
A value stored in the double data format requires 8\,bytes of space.
Thus, the data traffic is decreased by (512$-$4)$\times$8\,B\,=\,4,064\,B\,=\,3.97\,MB per sampling event, decreasing data traffic by 99.3\,\%.

%----------------------------------------------------------------------------------------

\section*{Summary}

This paper has presented a decentralized autonomous sensor fault detection strategy for structural health monitoring systems based on neural networks. 
Autonomous sensor fault detection has been realized by implementing neural networks into the sensor nodes.
The neural networks have been trained to predict expected sensor data to be compared to measured data, \colorbox{cyan}{in order to} detect sensor faults.
To verify the proposed approach, the SHM system has been installed on a test structure to be tested in laboratory experiments.
The structural integrity of the observed structure is monitored by calculating the natural frequency.
Simulated sensor faults have been detected reliably.
Through decentralized data processing, implemented into each sensor node, the data traffic has been reduced by 99.3\,\%.
In summary, it has been proven, that sensor fault detection using neural networks can improve the dependability and the accuracy of structural health monitoring systems.

%----------------------------------------------------------------------------------------

\bibliographystyle{unsrtnat}
\bibliography{literature}

\end{document}